---
title: "Biometry Ch. 6, section 6.7"
output: html_fragment
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<h3>Section 6.7 Graphic Methods</h3>

Figure 6.5 shows the cumulative normal distribution again (as in Fig. 6.3), with lines drawn down to equivalent deviates.
```{r}
cumulativecurve <- curve(pnorm(x, mean=0, sd=1),
      -4, 4,
      lty="solid")
```

This is the same code as earlier in [section 6.3](LINK HERE).
You can draw lines, we'll just do the one for 0.7 cumulative.
Remember that qnorm gets us from the cumulative distribution to the value.
Here are some examples of qnorm again.
```{r}
qnorm(0.7)
qnorm(0.3)
qnorm(1) #As the text says, this is Inf(infinity)
```
We'll use segments() to add arbitrary lines to the plot we made above.
This code will not draw anything if you don't have a plot open.

```{r}
plot(cumulativecurve,
     type="l")
segments(x0 = qnorm(0.7),
         y0 = 0,
         x1 = qnorm(0.7),
         y1 = 0.7)

segments(x0 = qnorm(0.7),
         y0 = 0.7,
         x1 = -4,
         y1 = 0.7)
```

You can generalize this by making a function.

```{r}
fig6.5.function <- function (cumulativepercent,
                             normaldeviateleft,
                             normaldeviateright){
  curve(pnorm(x, mean=0, sd=1),
        normaldeviateleft,
        normaldeviateright,
        lty="solid")
  
  segments(x0 = qnorm(cumulativepercent),
           y0 = 0,
           x1 = qnorm(cumulativepercent),
           y1 = cumulativepercent)
  
  segments(x0 = qnorm(cumulativepercent),
           y0 = cumulativepercent,
           x1 = normaldeviateleft,
           y1 = cumulativepercent)
}

fig6.5.function(0.7,
                -4,
                4)


fig6.5.function(0.3,
                -4,
                4)

```

On page 111, they begin discussing quantile-quantile plots (or QQ plots)

We'll make a Q-Q plot manually following box 6.2 to understand how they are built.
I. routine computer processed samples.
First we need to generate a small dataset.
```{r}
fakehouseflydata <- data.frame(
  "winglength" = rnorm(n = 1400,
                       mean = 45.5,
                       sd = 3.90)
)
```
Step 1.  order by increasing magnitude.

```{r message = FALSE}
library(dplyr)
```
```{r}
ordered_fakehouseflydata <- fakehouseflydata %>%
  arrange(winglength)
```

Step 2. Rank and compute p.
```{r}
ordered_fakehouseflydata$ranks <- rank(ordered_fakehouseflydata$winglength,
                                       ties.method = "first")
ordered_fakehouseflydata$p <- (ordered_fakehouseflydata$ranks - 0.5)/length(ordered_fakehouseflydata$winglength)
                                       
```

The rank() function has several methods to deal with ties.
Why did we use "first"?  Well, first let's look at the ranking options for ties and the order() function

```{r}
rank(c(4,3,6,2,2),
     ties.method = "first")

rank(c(4,3,6,2,2),
     ties.method = "last")

rank(c(4,3,6,2,2),
     ties.method = "average")

rank(c(4,3,6,2,2),
     ties.method = "random")

rank(c(4,3,6,2,2),
     ties.method = "min")

rank(c(4,3,6,2,2),
     ties.method = "max")
```
It looks like from step 2's description of how to deal with ties in i (the rank), we need something that will give us ties,
so first, last, and random are out.

R appears to use two nested order() functions.  [This description of how to make a QQ plot manually](https://rpubs.com/markheckmann/45771) points out that you can view the code.
I was a bit baffled about why two order functions, but [this post](http://stackoverflow.com/questions/2315601/understanding-the-order-function) and [its linked quote](https://stat.ethz.ch/pipermail/r-help/2001-November/016802.html) helped me understand it. 
The first order gives you the row indices that would put the numbers from smallest to largest (in ascending order).
The nesting of order orders those new numbers (the 4, 5, 2, 1,3) by the same procedure.
In other words, it gives the row number of the vector of row numbers that puts the original vector from smallest to largest.
It is not intuitive why you would use this versus rank().  Rank just goes ahead and ranks from smallest to largest,
so you can put that value directly into a column.
Not sure if you can put the order(order()) into a column.

```{r}
order(c(4,3,6,2,2))
order(order(c(4,3,6,2,2)))
```


However, [this description of how to make a QQ plot manually](https://rpubs.com/markheckmann/45771) specifies that you don't want ties.  So there is more than one way to do this.


Step 3. Computed normal quantiles for each p.
```{r}
ordered_fakehouseflydata$normal_quantile <- qnorm(ordered_fakehouseflydata$p,
                                                  mean = 0, 
                                                  sd = 1)
```

Step 4. plot
```{r}
plot(y = ordered_fakehouseflydata$normal_quantile,
     x = ordered_fakehouseflydata$winglength)
```

More guidance on making these from scratch [here](https://rpubs.com/markheckmann/45771)

http://stats.stackexchange.com/questions/230226/how-to-manually-make-a-qq-plot-of-p-value-against-uniform-distribution
A different way, see how it compares.
More background http://onlinestatbook.com/2/advanced_graphs/q-q_plots.html

```{r}
stats:::qqnorm.default
x <- qnorm(ppoints(n))[order(order(y))]
```

The rankits in box 6.2 are done this way.

 Look up Q-Q calculation in R to see how it deals with small sample sizes.
 
Then how to do them in R automatically.
 
