---
title: "Biometry Ch. 6, section 6.7"
output: html_fragment
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<h3>Section 6.7 Graphic Methods</h3>

Figure 6.5 shows the cumulative normal distribution again (as in Fig. 6.3), with lines drawn down to equivalent deviates.
```{r}
cumulativecurve <- curve(pnorm(x, mean=0, sd=1),
      -4, 4,
      lty="solid")
```

This is the same code as earlier in [section 6.3](LINK HERE).
You can draw lines, we'll just do the one for 0.7 cumulative.
Remember that qnorm gets us from the cumulative distribution to the value.
Here are some examples of qnorm again.
```{r}
qnorm(0.7)
qnorm(0.3)
qnorm(1)    #As the text says, this is Inf(infinity)
qnorm(0)    #Ditto
qnorm(-0.7) #Remember because you are inputting a vector of probabilities it must be between 0 and 1.
```
We'll use segments() to add arbitrary lines to the plot we made above.
This code will not draw anything if you don't have a plot open.

```{r}
plot(cumulativecurve,
     type="l")
segments(x0 = qnorm(0.7),
         y0 = 0,
         x1 = qnorm(0.7),
         y1 = 0.7)

segments(x0 = qnorm(0.7),
         y0 = 0.7,
         x1 = -4,
         y1 = 0.7)
```

You can generalize this by making a function.

```{r}
fig6.5.function <- function (cumulativepercent,
                             normaldeviateleft,
                             normaldeviateright){
  curve(pnorm(x, mean=0, sd=1),
        normaldeviateleft,
        normaldeviateright,
        lty="solid")
  
  segments(x0 = qnorm(cumulativepercent),
           y0 = 0,
           x1 = qnorm(cumulativepercent),
           y1 = cumulativepercent)
  
  segments(x0 = qnorm(cumulativepercent),
           y0 = cumulativepercent,
           x1 = normaldeviateleft,
           y1 = cumulativepercent)
}

fig6.5.function(0.7,
                -4,
                4)


fig6.5.function(0.3,
                -4,
                4)

```

On page 111, they begin discussing quantile-quantile plots (or QQ plots)

We'll make a Q-Q plot manually following box 6.2 to understand how they are built.
I. routine computer processed samples.
First we need to generate a small dataset.
```{r}
set.seed(24652498)  #set the random seed so we get the same results each time.
fakehouseflydata <- data.frame(
  "winglength" = rnorm(n = 1400,
                       mean = 45.5,
                       sd = 3.90)
)
```
Step 1.  order by increasing magnitude.

```{r message = FALSE}
library(dplyr)
```
```{r}
ordered_fakehouseflydata <- fakehouseflydata %>%
  arrange(winglength)
```

Step 2. Rank and compute p.
The rank() function has several methods to deal with ties.
First let's look at the ranking options for ties.

```{r}
rank(c(4,3,6,2,2),
     ties.method = "first")

rank(c(4,3,6,2,2),
     ties.method = "last")

rank(c(4,3,6,2,2),
     ties.method = "average")

rank(c(4,3,6,2,2),
     ties.method = "random")

rank(c(4,3,6,2,2),
     ties.method = "min")

rank(c(4,3,6,2,2),
     ties.method = "max")
```
It looks like from step 2's description of how to deal with ties in i (the rank), we need something that will give us ties,
so first, last, and random are out.  Let's do average.

```{r}
ordered_fakehouseflydata$i <- rank(ordered_fakehouseflydata$winglength,
                                       ties.method = "average")
ordered_fakehouseflydata$p <- (ordered_fakehouseflydata$i - 0.5)/length(ordered_fakehouseflydata$winglength)
                                       
```

The rank order divided by n (sample size) will result in the first observation being small (1-1/2)/n but above 0 and the last observation being closer to 1.  These are [percentiles](https://books.google.com/books?id=Xj3SBQAAQBAJ&pg=PA108&lpg=PA108&dq=p+%3D+(i+-1/2)/n+quantile&source=bl&ots=w5-cCqQMJE&sig=Ahu8gN1tFHPMCERhqghUv0qjnaI&hl=en&sa=X&ved=0ahUKEwijsoabpLPSAhUBHWMKHWrQAmsQ6AEIJTAC#v=onepage&q=p%20%3D%20(i%20-1%2F2)%2Fn%20quantile&f=false) if you multiplied them by 100.  R also has a quantile() function to make sample quantiles (aka percentiles at various probabilities).  Type 5 uses the same equation as we did to make the variable "p".  You can see what it does below.  The difference appears to be that quantile() tells you what values exist at given percentiles (the variables you put into the probs argument.)  In our case, we are calculating in what percentile the sample is.  So we wouldn't use quantile() in this case despite its appealing name relative to step 3 below.

```{r}
quantile(x = ordered_fakehouseflydata$winglength,
         type = 5)
```


Step 3. Computed normal quantiles for each p.  Here we take the "p" column and input it into qnorm.  qnorm() asks for a vector of probabilities (see ?qnorm).  But our "p" is calculated in a way like the sample quantiles/percentiles in quantile().  [Here is an explanation.](http://stats.stackexchange.com/questions/132952/is-calculating-a-percentile-the-same-as-evaluating-a-cumulative-density-function) and [this one too](https://www.fs.fed.us/rm/pubs_int/int_gtr367/12comp.html)
```{r}
ordered_fakehouseflydata$normal_quantile <- qnorm(ordered_fakehouseflydata$p,
                                                  mean = 0, 
                                                  sd = 1)
```

Step 4. Plot.
You can use R's qqline to "draw the line with a ruler" as they suggest.  
R uses the opposite axes compared to the book so we have to use the datax=TRUE argument to get it to match.
```{r}
plot(y = ordered_fakehouseflydata$normal_quantile,
     x = ordered_fakehouseflydata$winglength,
     xlab = "Housefly wing lengths (Y)",
     ylab = "Normal quantiles")
qqline(y=ordered_fakehouseflydata$winglength,
       datax=TRUE)

qqnorm(y = ordered_fakehouseflydata$winglength,
       datax = TRUE)
qqline(y = ordered_fakehouseflydata$winglength,
       datax = TRUE)



plot(y = ordered_fakehouseflydata$normal_quantilei,
     x = ordered_fakehouseflydata$winglength,
     xlab = "Housefly wing lengths (Y)",
     ylab = "Normal quantiles")
qqline(y=ordered_fakehouseflydata$winglength,
       datax=TRUE)
```
From this you can see that what the book calls normal quantiles, R calls theoretical quantiles.
More guidance on making these from scratch [here](https://rpubs.com/markheckmann/45771)

http://stats.stackexchange.com/questions/230226/how-to-manually-make-a-qq-plot-of-p-value-against-uniform-distribution
A different way, see how it compares.
More background http://onlinestatbook.com/2/advanced_graphs/q-q_plots.html


R can do this all for you automatically, as you can see from using qqnorm and qqline.  qqplot() will do both together automatically.  How does R's function stack up with our code?

First, R appears to use two nested order() functions instead of rank.  [This description of how to make a QQ plot manually](https://rpubs.com/markheckmann/45771) points out that you can view the code.
I was a bit baffled about why two order functions, but [this post](http://stackoverflow.com/questions/2315601/understanding-the-order-function) and [its linked quote](https://stat.ethz.ch/pipermail/r-help/2001-November/016802.html) helped me understand it. 
The first order gives you the row indices that would put the numbers from smallest to largest (in ascending order).
The nesting of order orders those new numbers (the 4, 5, 2, 1,3) by the same procedure.  Order's result is always row indices.  Rank's result is always what rank the value is (1, 2, 3, etc), not its row index.  [Rank is equivalent to order(order).](http://stackoverflow.com/questions/38968911/why-does-orderorderx-equal-rankx-in-r?noredirect=1&lq=1)  Essentially when you order the order indices (i.e. get their row indices) it gives you the rank.


```{r}
order(c(4,3,6,2,2))
order(order(c(4,3,6,2,2)))
```


Additionally, [this description of how to make a QQ plot manually](https://rpubs.com/markheckmann/45771) specifies that you don't want ties.  So there is more than one way to do this [see also the many ways to calculate quantiles in the quantile() function)].


```{r}
stats:::qqnorm.default
x <- qnorm(ppoints(n))[order(order(y))]
```
Specifically, R uses qnorm().  Qnorm generates quantiles for the normal distribution.

An example of how to interpret this is [here](http://data.library.virginia.edu/understanding-q-q-plots/)

Part II of box 6.2 shows up to do a QQ plot with a frequency distribution.  The main difference here is in the calculations.  If we wanted to make a QQ plot in R, we just take the class marks and do qqplot() on them.  When calculating by hand, we need the cumulative frequencies. Let's [add up the frequency (f) column](http://stackoverflow.com/questions/33038240/add-row-value-to-previous-row-value-in-r).

df1$newColumn <- with(df1, c(f[1],f[-1]+f[-nrow(df1)]))

The rankits in box 6.2 are done this way.

 Look up Q-Q calculation in R to see how it deals with small sample sizes.
 
Then how to do them in R automatically.
 
