---
title: "Biometry Ch. 6, section 6.7"
output: html_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<h3>Section 6.7 Graphic Methods</h3>

Figure 6.5 shows the cumulative normal distribution again (as in Fig. 6.3), with lines drawn down to equivalent deviates.
```{r}
cumulativecurve <- curve(pnorm(x, mean=0, sd=1),
      -4, 4,
      lty="solid")
```

This is the same code as earlier in [section 6.3](LINK HERE).
You can draw lines, we'll just do the one for 0.7 cumulative.
Remember that qnorm gets us from the cumulative distribution to the value.
Here are some examples of qnorm again.
```{r}
qnorm(0.7)
qnorm(0.3)
qnorm(1) #As the text says, this is Inf(infinity)
```
We'll use segments() to add arbitrary lines to the plot we made above.
This code will not draw anything if you don't have a plot open.

```{r}
plot(cumulativecurve,
     type="l")
segments(x0 = qnorm(0.7),
         y0 = 0,
         x1 = qnorm(0.7),
         y1 = 0.7)

segments(x0 = qnorm(0.7),
         y0 = 0.7,
         x1 = -4,
         y1 = 0.7)
```

You can generalize this by making a function.

```{r}
fig6.5.function <- function (cumulativepercent,
                             normaldeviateleft,
                             normaldeviateright){
  curve(pnorm(x, mean=0, sd=1),
        normaldeviateleft,
        normaldeviateright,
        lty="solid")
  
  segments(x0 = qnorm(cumulativepercent),
           y0 = 0,
           x1 = qnorm(cumulativepercent),
           y1 = cumulativepercent)
  
  segments(x0 = qnorm(cumulativepercent),
           y0 = cumulativepercent,
           x1 = normaldeviateleft,
           y1 = cumulativepercent)
}

fig6.5.function(0.7,
                -4,
                4)


fig6.5.function(0.3,
                -4,
                4)

```

On page 111, they begin discussing quantile-quantile plots (or QQ plots)

Make one manually.
Following box 6.2 to manually make a Q-Q plot to understand how they are built.
I. routine computer processed samples.
First we need to generate a small dataset.
```{r}
fakehouseflydata <- data.frame(
  "winglength" = rnorm(n = 1400,
                       mean = 45.5,
                       sd = 3.90)
)
```
Step 1.  order by increasing magnitude.

```{r message = FALSE}
library(dplyr)
```
```{r}
ordered_fakehouseflydata <- fakehouseflydata %>%
  arrange(winglength)
```

Step 2. Rank and compute p.
```{r}
ordered_fakehouseflydata$ranks <- rank(ordered_fakehouseflydata$winglength,
                                       ties.method = "first")
ordered_fakehouseflydata$p <- (ordered_fakehouseflydata$ranks - 0.5)/length(ordered_fakehouseflydata$winglength)
                                       
```

Then how to do them in R automatically.
Which ranking method should you choose?

```{r}
rank(c(1,2,3,4,4,5),
     ties.method = "first")

rank(c(1,2,3,4,4,5),
     ties.method = "last")

rank(c(1,2,3,4,4,5),
     ties.method = "average")

rank(c(1,2,3,4,4,5),
     ties.method = "random")

rank(c(1,2,3,4,4,5),
     ties.method = "min")

rank(c(1,2,3,4,4,5),
     ties.method = "max")
```

Step 3. Computed normal quantiles for each p.
```{r}
ordered_fakehouseflydata$normal_quantile <- qnorm(ordered_fakehouseflydata$p,
                                                  mean = 0, 
                                                  sd = 1)
```

Step 4. plot
```{r}
plot(y = ordered_fakehouseflydata$normal_quantile,
     x = ordered_fakehouseflydata$winglength)
```

More guidance on making these from scratch [here](https://rpubs.com/markheckmann/45771)

http://stats.stackexchange.com/questions/230226/how-to-manually-make-a-qq-plot-of-p-value-against-uniform-distribution
A different way, see how it compares.
More background http://onlinestatbook.com/2/advanced_graphs/q-q_plots.html

The rankits in box 6.2 are done this way.

 Look up Q-Q calculation in R to see how it deals with small sample sizes.
 
 
